---
title: "Using the SMP Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the SMP Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message=FALSE}
library(SMP)
seed <- 20010416
```

This vignette demonstrates the sequential mixture prior (SMP) functions for binary and continuous endpoints using simulated data. The examples use `pi2_prior = "MAP"`; the same function calls can use `pi2_prior = "pool"` so that the prior for case when both nonconcurrent and historical controls are congruent is constructed from the posterior based on the pooled nonconcurrent and historical controls.

# Binary Endpoint

## Data Simulation

We first simulate a binary endpoint with treatment, concurrent control, nonconcurrent control, and historical control data.

```{r binary-simulation}
set.seed(seed)

binary_truth <- c(theta_t = 0.55, theta_c = 0.40, theta_n = 0.40, theta_h = 0.2)
binary_n <- c(nt = 200, nc = 100, nn = 200, nh = 200)

binary_x <- c(
  xt = rbinom(1, binary_n[["nt"]], binary_truth[["theta_t"]]),
  xc = rbinom(1, binary_n[["nc"]], binary_truth[["theta_c"]]),
  xn = rbinom(1, binary_n[["nn"]], binary_truth[["theta_n"]]),
  xh = rbinom(1, binary_n[["nh"]], binary_truth[["theta_h"]])
)

binary_data <- data.frame(
  source = c("Treatment", "Concurrent", "Nonconcurrent", "Historical"),
  n = unname(binary_n[c("nt", "nc", "nn", "nh")]),
  responders = unname(binary_x[c("xt", "xc", "xn", "xh")])
)
binary_data$response_rate <- binary_data$responders / binary_data$n
binary_data
```

## SMP

We fit the binary SMP model using a clinically meaningful difference of 0.10.

```{r binary-fit}
binary_fit <- smp_binary(
  nt = binary_n[["nt"]], xt = binary_x[["xt"]],
  nc = binary_n[["nc"]], xc = binary_x[["xc"]],
  nn = binary_n[["nn"]], xn = binary_x[["xn"]],
  nh = binary_n[["nh"]], xh = binary_x[["xh"]],
  delta = 0.10,
  n_samples = 5000,
  pi2_prior = "MAP",
  seed = seed
)
```

## Output

The fitted object contains posterior samples and the borrowing weights.

```{r binary-output}
binary_summary <- summary(binary_fit)
binary_summary$posterior
binary_summary$prob_superiority
transform(binary_summary$weights, weight = round(weight, 2))
```


# Continuous Endpoint

## Data Simulation

We next simulate a normally distributed endpoint. The treatment arm has a higher mean than the concurrent control, and the external controls are close to the concurrent control.

```{r continuous-simulation}
set.seed(seed)

continuous_n <- c(nt = 200, nc = 100, nn = 200, nh = 200)
continuous_truth <- c(mu_t = 0.60, mu_c = 0.00, mu_n = 0.00, mu_h = 1.00)
sigma <- 2

yt <- rnorm(continuous_n[["nt"]], mean = continuous_truth[["mu_t"]], sd = sigma)
yc <- rnorm(continuous_n[["nc"]], mean = continuous_truth[["mu_c"]], sd = sigma)
yn <- rnorm(continuous_n[["nn"]], mean = continuous_truth[["mu_n"]], sd = sigma)
yh <- rnorm(continuous_n[["nh"]], mean = continuous_truth[["mu_h"]], sd = sigma)

continuous_data <- data.frame(
  source = c("Treatment", "Concurrent", "Nonconcurrent", "Historical"),
  n = unname(continuous_n[c("nt", "nc", "nn", "nh")]),
  mean = c(mean(yt), mean(yc), mean(yn), mean(yh)),
  sd = c(sd(yt), sd(yc), sd(yn), sd(yh))
)
continuous_data
```

## SMP

We fit the continuous SMP model using a clinically meaningful mean difference of 0.60.

```{r continuous-fit}
continuous_fit <- smp_continuous(
  nt = length(yt), yt = yt,
  nc = length(yc), yc = yc,
  nn = length(yn), yn = yn,
  nh = length(yh), yh = yh,
  delta = 0.60,
  n_samples = 5000,
  pi2_prior = "MAP",
  seed = seed
)
```

## Output

```{r continuous-output}
continuous_summary <- summary(continuous_fit)
continuous_summary$posterior
continuous_summary$prob_superiority
transform(continuous_summary$weights, weight = round(weight, 2))
```
